---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  fig.retina = 3 
)
```

# {ggspc}

<!-- badges: start -->
<!-- badges: end -->

The goal of {ggspc} is to provide custom 'Stat', 'Geom' and 'theme' definitions for 'SoilProfileCollection' object compatibility with 'ggplot2'.


## Installation

You can install the development version of {ggspc} like so:

``` r
remotes::install_github("brownag/ggspc")
```

## Example

This a example shows how to solve the common problem of plotting variables contained in a `SoilProfileCollection` with `ggplot2::ggplot()`

```{r example}
library(aqp)
library(ggspc)
library(ggplot2)

data(loafercreek, package = "soilDB")
GHL(loafercreek) <- "genhz"
```

### Basics

This is a demonstration of what is possible with a simple `fortify(<SPC>)` method defined. The "fortify" method makes it such that names from horizon and site slots of the SPC can be used in `ggplot()` aesthetics via `aes()`. 

```{r basic-spc-ggplot}
# site v.s. site level
ggplot(loafercreek, aes(earthcovkind1, slope_field)) +
  geom_boxplot(na.rm = TRUE)
```

```{r basic-spc-ggplot2}
# site v.s. horizon level
ggplot(loafercreek, aes(hillslopeprof, clay)) +
  geom_boxplot(na.rm = TRUE)
```

```{r basic-spc-ggplot3}
# horizon v.s. horizon level
ggplot(loafercreek, aes(clay, genhz)) +
  geom_boxplot(na.rm = TRUE)
```

### Depth Weighted Aggregation (`stat_depth_weighted()`)

The `stat_depth_weighted()` function is a specialized {ggplot2} statistic intended for calculation of depth-weighted values for horizon data in a `SoilProfileCollection`. The default uses a constant interval `from=0` `to=200` (centimeters), but the intervals of interest may alternately be specified as site-level column names (unquoted), and therefore may vary between profiles.

Currently, `stat_depth_weighted()` only supports the "point" geometry type, but in future "boxplot" and others may be supported.

```{r stat-depth-weighted-spc}
# default y aesthetic is the profile_id(<SPC>)
ggplot(loafercreek[1:10], aes(clay)) +
  stat_depth_weighted(na.rm = TRUE)
```

```{r stat-depth-weighted-spc2}
# can use alternate y aesthetic, e.g. hillslopeprof
ggplot(loafercreek, aes(clay, hillslopeprof)) +
  stat_depth_weighted(na.rm = TRUE)
```

```{r stat-depth-weighted-spc3}
# continuous y axes works too (horizon v.s. horizon)
ggplot(loafercreek, aes(clay, sand)) +
  stat_depth_weighted(na.rm = TRUE)
```

```{r stat-depth-weighted-spc4}
# continuous y (horizon v.s. site)
ggplot(loafercreek, aes(clay, slope_field)) +
  stat_depth_weighted(na.rm = TRUE)
```

```{r stat-depth-weighted-spc5}
# can combine with typical ggplot geoms (0-200cm mean over boxplots)
ggplot(loafercreek, aes(clay, hillslopeprof)) +
  geom_boxplot(na.rm = TRUE) + 
  stat_depth_weighted(na.rm = TRUE, col = "red", pch = 17, cex = 3)
```

```{r stat-depth-weighted-spc6}
# use site-level columns for profile-specific intervals (e.g. PSCS)
ggplot() +
  stat_depth_weighted(
    loafercreek,
    aes(clay, hillslopeprof),
    na.rm = TRUE,
    from = psctopdepth,
    to = pscbotdepth
  )
```

## Problem

Currently the stat_* function extracts metadata after the `fortify()` method is called by accessing a "last SoilProfileCollection" cached in a special environment `ggspc.env` exported by the package.

This causes problems for the times when the data are defined in the `ggplot()` call, such as the following unexpected behavior. This follows from the hacky-ness of the aforementioned caching side effect/dependency. This can be "fixed" with an explicit reference to the dataset (i.e. `x + stat_depth_weighted(loafercreek)`). 

Perhaps there is a better, more canonical ggplot2 way to implement this I have still not figured out. As an alternative to overwriting a last SPC object in a shared package environment, I return the SPC as an attribute of the `ggplot()` result, but I don't know that it is possible to expose that attribute to the `stat_*` functions or `ggproto` classes by any standard method.


```{r, error = TRUE}
data(loafercreek, package = "soilDB")
data(gopheridge, package = "soilDB")

x <- ggplot(loafercreek, aes(clay, hillslopeprof))
x <- ggplot(gopheridge, aes(clay, hillslopeprof))

x + geom_boxplot() # works
y + geom_boxplot() # works

x + stat_depth_weighted() # error
x + stat_depth_weighted(loafercreek) # works
y + stat_depth_weighted() # error
x + stat_depth_weighted(gopheridge) # works
```
